---
title: "AdapVal_CRC Example"
author: "Lindsay Collin"
date: "December 23, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyverse)
library(sas7bdat)
library(binom)
```

```{r CRC validation data}
dat<-read.sas7bdat("analysis_CRC.sas7bdat")
```

## Data Cleaning
select only relevant variables
Create new id variable
Order study population by study time (min)
order outcome by event time

```{r cleaning}
myvars<-c("OP_DATO", "d_recurrence","event","d_event","recidiv_coms","recidivdat_coms","recurrence")
dat1<-dat[myvars]
summary(dat1)
table(dat1$recurrence, dat1$recidiv_coms)
```
```{r cleaning1}
dat1$studytime<-dat1$OP_DATO-min(dat1$OP_DATO)
dat1$rtime<-dat1$d_recurrence-min(dat1$OP_DATO)
dat1$R<-ifelse(dat1$rtime>=0,1,NA)
dat1$S<-1
dat2<-dat1[with(dat1,order(studytime)),]
dat2$S<-cumsum(dat2$S)
dat3<-dat2[with(dat2,order(rtime)),]
dat3$R<-cumsum(dat3$R)
dat3$id<-1:nrow(dat3)
```
## Adaptive Validation--T1
Create indicator for if someone has been validated (set to 0)
subset after first 5 cases, make note of those who may go on to develop Recurrence (exclude)

```{r Adap Val}
dat3$T<-0
set.seed(20191506)
dat3 <-dat3[order(dat3$studytime),]
dat3$R[is.na(dat3$R)]<-0
dat4<- dat3 %>% filter(studytime<=rtime[R==5])

#PPV
samp1<-dat4 %>% filter(R %in% 1:5)
table(samp1$recidiv_coms)
#NPV
samp2<-dat4 %>% filter(R==0)
samp2<-samp2%>%sample_n(5, replace=FALSE)
table(samp2$recidiv_coms)


```
## Adaptive Validation--T2
Create indicator for if someone has been validated (set to 0)
subset after first 5 cases, make note of those who may go on to develop Recurrence
```{r T2}
#identify validation set in original dataset

dat3$T<-ifelse(dat3$id %in% samp2$id,1,dat3$T)
table(dat3$T)

dat5<- dat3 %>% filter(studytime<=rtime[R==10])
dat5<-subset(dat5, T==0)
#PPV
samp3<-dat5 %>% filter(R %in% 6:10)
table(samp3$recidiv_coms)
#NPV
samp4<-dat5 %>% filter(R==0)
samp4<-samp4%>%sample_n(5, replace=FALSE)
table(samp4$recidiv_coms)

```

## Adaptive Validation--T3
Create indicator for if someone has been validated (set to 0)
subset after first 5 cases, make note of those who may go on to develop Recurrence
```{r T3}
dat3$T<-ifelse(dat3$id %in% samp4$id,2,dat3$T)
table(dat3$T)

dat6<- dat3 %>% filter(studytime<=rtime[R==15])
dat6<-subset(dat6, T==0)
#PPV
samp6<-dat6 %>% filter(R %in% 11:15)
table(samp6$recidiv_coms)
#NPV
samp7<-dat6 %>% filter(R==0)
samp7<-samp7%>%sample_n(5, replace=FALSE)
table(samp7$recidiv_coms)
```
## Adaptive Validation--T4
Create indicator for if someone has been validated (set to 0)
subset after first 5 cases, make note of those who may go on to develop Recurrence

```{r T4}
dat3$T<-ifelse(dat3$id %in% samp7$id,3,dat3$T)
table(dat3$T)
dat7<- dat3 %>% filter(studytime<=rtime[R==20])
dat7<-subset(dat7, T==0)
#PPV
samp8<-dat7 %>% filter(R %in% 16:20)
table(samp8$recidiv_coms)
#NPV
samp9<-dat7 %>% filter(R==0)
samp9<-samp9%>%sample_n(5, replace=FALSE)
table(samp9$recidiv_coms)
```

## Adaptive Validation--T5
Create indicator for if someone has been validated (set to 0)
subset after first 5 cases, make note of those who may go on to develop Recurrence
```{r T5}
dat3$T<-ifelse(dat3$id %in% samp9$id,4,dat3$T)
table(dat3$T)
dat8<- dat3 %>% filter(studytime<=rtime[R==25])
dat8<-subset(dat8, T==0)
#PPV
samp10<-dat8 %>% filter(R %in% 21:25)
table(samp10$recidiv_coms)
#NPV
samp11<-dat8 %>% filter(R==0)
samp11<-samp11%>%sample_n(5, replace=FALSE)
table(samp11$recidiv_coms)
```
## Adaptive Validation--T6
Create indicator for if someone has been validated (set to 0)
subset after first 5 cases, make note of those who may go on to develop Recurrence
```{r T6}
dat3$T<-ifelse(dat3$id %in% samp11$id,5,dat3$T)
table(dat3$T)
dat9<- dat3 %>% filter(studytime<=rtime[R==30])
dat9<-subset(dat9, T==0)
#PPV
samp12<-dat9 %>% filter(R %in% 26:30)
table(samp12$recidiv_coms)
#NPV
samp13<-dat9 %>% filter(R==0)
samp13<-samp13%>%sample_n(5, replace=FALSE)
table(samp13$recidiv_coms)

```

## Adaptive Validation--T7
Create indicator for if someone has been validated (set to 0)
subset after first 5 cases, make note of those who may go on to develop Recurrence
```{r}
dat3$T<-ifelse(dat3$id %in% samp13$id,6,dat3$T)
table(dat3$T)
dat10<- dat3 %>% filter(studytime<=rtime[R==35])
dat10<-subset(dat10, T==0)
#PPV
samp14<-dat10 %>% filter(R %in% 31:35)
table(samp14$recidiv_coms)
#NPV
samp15<-dat10 %>% filter(R==0)
samp15<-samp15%>%sample_n(5, replace=FALSE)
table(samp15$recidiv_coms)

```
## Adaptive Validation--T8
Create indicator for if someone has been validated (set to 0)
subset after first 5 cases, make note of those who may go on to develop Recurrence
```{r}
dat3$T<-ifelse(dat3$id %in% samp15$id,7,dat3$T)
table(dat3$T)
dat11<- dat3 %>% filter(studytime<=rtime[R==40])
dat11<-subset(dat11, T==0)
#PPV
samp16<-dat11 %>% filter(R %in% 36:40)
table(samp16$recidiv_coms)
#NPV
samp17<-dat11 %>% filter(R==0)
samp17<-samp17%>%sample_n(5, replace=FALSE)
table(samp17$recidiv_coms)
```
## Adaptive Validation--T9
Create indicator for if someone has been validated (set to 0)
subset after first 5 cases, make note of those who may go on to develop Recurrence
```{r}
dat3$T<-ifelse(dat3$id %in% samp17$id,8,dat3$T)
table(dat3$T)
dat12<- dat3 %>% filter(studytime<=rtime[R==45])
dat12<-subset(dat12, T==0)
#PPV
samp18<-dat12 %>% filter(R %in% 41:45)
table(samp18$recidiv_coms)
#NPV
samp19<-dat12 %>% filter(R==0)
samp19<-samp19%>%sample_n(5, replace=FALSE)
table(samp19$recidiv_coms)
```
## Adaptive Validation--T10
Create indicator for if someone has been validated (set to 0)
subset after first 5 cases, make note of those who may go on to develop Recurrence
```{r}
dat3$T<-ifelse(dat3$id %in% samp19$id,9,dat3$T)
table(dat3$T)
dat13<- dat3 %>% filter(studytime<=rtime[R==50])
dat13<-subset(dat13, T==0)
#PPV
samp20<-dat13 %>% filter(R %in% 46:50)
table(samp20$recidiv_coms)
#NPV
samp21<-dat13 %>% filter(R==0)
samp21<-samp21%>%sample_n(5, replace=FALSE)
table(samp21$recidiv_coms)
```
## Adaptive Validation--T11
Create indicator for if someone has been validated (set to 0)
subset after first 5 cases, make note of those who may go on to develop Recurrence
```{r}
dat3$T<-ifelse(dat3$id %in% samp21$id,10,dat3$T)
table(dat3$T)
dat14<- dat3 %>% filter(studytime<=rtime[R==55])
dat14<-subset(dat14, T==0)
#PPV
samp22<-dat14 %>% filter(R %in% 51:55)
table(samp22$recidiv_coms)
#NPV
samp23<-dat14 %>% filter(R==0)
samp23<-samp23%>%sample_n(5, replace=FALSE)
table(samp23$recidiv_coms)
```

## Adaptive Validation--T12
Create indicator for if someone has been validated (set to 0)
subset after first 5 cases, make note of those who may go on to develop Recurrence
```{r}
dat3$T<-ifelse(dat3$id %in% samp23$id,11,dat3$T)
table(dat3$T)
dat15<- dat3 %>% filter(studytime<=rtime[R==60])
dat15<-subset(dat15, T==0)
#PPV
samp24<-dat15 %>% filter(R %in% 56:60)
table(samp24$recidiv_coms)
#NPV
samp25<-dat15 %>% filter(R==0)
samp25<-samp25%>%sample_n(5, replace=FALSE)
table(samp25$recidiv_coms)
```

## Adaptive Validation--T12
Create indicator for if someone has been validated (set to 0)
subset after first 5 cases, make note of those who may go on to develop Recurrence
```{r}
dat3$T<-ifelse(dat3$id %in% samp25$id,12,dat3$T)
table(dat3$T)
dat16<- dat3 %>% filter(studytime<=rtime[R==65])
dat16<-subset(dat16, T==0)
#PPV
samp26<-dat16 %>% filter(R %in% 61:65)
table(samp26$recidiv_coms)
#NPV
samp27<-dat16 %>% filter(R==0)
samp27<-samp27%>%sample_n(5, replace=FALSE)
table(samp27$recidiv_coms)
```
## Adaptive Validation--T13
Create indicator for if someone has been validated (set to 0)
subset after first 5 cases, make note of those who may go on to develop Recurrence

```{r}
dat3$T<-ifelse(dat3$id %in% samp27$id,13,dat3$T)
table(dat3$T)
dat17<- dat3 %>% filter(studytime<=rtime[R==70])
dat17<-subset(dat17, T==0)
#PPV
samp28<-dat17 %>% filter(R %in% 66:70)
table(samp28$recidiv_coms)
#NPV
samp29<-dat17 %>% filter(R==0)
samp29<-samp29%>%sample_n(5, replace=FALSE)
table(samp29$recidiv_coms)
```

